{\fontfamily{ptm}\selectfont A Parameter Free Accelerated Proximal Gradient Method Without Restarting }

### 2024-12-12 Notebook LaTEX/Momentum Accelerations/paper3.tex
>w<, a lot of changes

\section{Preliminaries}
    Throughout Section 1, 2, 3, we assume the problem of types $F = f + g$ where $f: \RR \rightarrow \RR$ is $L$ Lipschitz smooth and $\mu \ge 0$ strongly convex and $g: Q \rightarrow \overline \RR$ is convex. 
    We consider optimization problem of the form
    \begin{align*}
        \min_x \left\lbrace
            F(x) \defeq f(x) + g(x)
        \right\rbrace. 
    \end{align*}
    To start we define the following quantities
    \begin{enumerate}
        \item The proximal gradient model function: 
        $$
            \widetilde{\mathcal M}^{L^{-1}}
            (x; y) \defeq
            g(x) + f(y) + \langle \nabla f(y), x - y\rangle 
            + \frac{L}{2}\Vert x - y\Vert^2. 
        $$
        \item The proximal point model function: 
        $$
            \mathcal M^{L^{-1}}(x; y) := F(x) + \frac{L}{2}\Vert x - y\Vert^2
        $$
        \item The proximal gradient operator $T_L = [I + L^{-1}\partial g]\circ [I - L^{-1}\nabla f]$. Since $L$ is fixed throughout, the subscript may be omitted and become $T$ instead to make the notations simpler. 
    \end{enumerate}
    To state the following lemma, we define the Bregman divergence of $f$ to be 
    \begin{align*}
        D_f(x, y): \RR^n \times \RR^n \rightarrow \RR 
        \defeq f(x) - f(y) - \langle \nabla f(x), x - y\rangle. 
    \end{align*}

@@

\section{Stepwise formulation of weak accelerated proximal gradient}
    In this section, we introduce the following new quantities and their notations. 
    It's there to make the notations simpler and easier to follow for the proofs that will come later. 
    The algorithm is parameterized by the sequence $y_k, x_k$. 
    We define for all $k \ge 0$ the following quantities: 
    \begin{assumption}
        Choose any integer $k\ge 0$. 
        Given $x_k, y_k, v_k$, we define the following quantities
        \begin{align}
            g_k &\defeq L(y_k - T_L y_k), 
            \label{eqn:grad-map}
            \\
            l_F(x; y_k) &\defeq F(T_Ly_k) + \langle g_k, x - y_k\rangle + \frac{1}{2L}\Vert g_k\Vert^2, 
            \label{eqn:lower-linearization}
            \\
            \epsilon_{k} &\defeq F(x_k) - l_F(x_k; y_k), 
            \label{eqn:regret}
        \end{align}
        Observe that by convexity of $F$, $\epsilon_k \ge 0$ always. 
    \end{assumption}

    \begin{definition}[Stepwise weak accelerated proximal gradient]\label{def:stepwise-wapg}\;\\
        Let $0 \le \mu \le L$ be the strong convexity and Lipschitz smoothness parameter of $f$. 
        Given iterates $(v_k, x_k)$, or equivalently $(y_k, x_k)$, $(x_k, y_k)$, any $\alpha_k \ge 0, \gamma \ge 0$, the algorithm generates scalar $\hat \gamma$, and vectors $y_k, v_{k + 1}, x_{k + 1}$ by equalities: 
        \begin{align*}
            \hat \gamma &:= (1 - \alpha_k)\gamma + \mu \alpha_k, 
            \\
            y_k &= 
            (\gamma + \alpha_k \mu)^{-1}(\alpha_k \gamma v_k + \hat\gamma x_k), 
            \\
            g_k &= \mathcal G_L y_k, 
            \\
            v_{k + 1} &= \hat\gamma^{-1}
            (\gamma(1 - \alpha_k) v_k - \alpha_k g_k + \mu \alpha_k y_k), 
            \\
            x_{k + 1} &= T_L y_k. 
        \end{align*}
    \end{definition}
    \begin{observation}\label{obs:stepwise-wapg}
        Let's observe Definition \ref{def:stepwise-wapg} closely. 
        We make two crucial observations here. 
        With the equalities, we define the following quantities associate with it. 
        Fix any integer $k \ge 0$. 
        To verify, we use the fundamental proximal gradient inequality. 
        Next, observe that we have $x_k, y_k, v_k$ lie on the same line because they have 
        \begin{align*}
            (Q1): 
            y_k - v_k &= 
            \frac{\hat \gamma}{\alpha_k \gamma}(x_k - y_k),
            \\
            (Q2): 
            y_k - x_k &= 
            \frac{\alpha_k \gamma}{\gamma + \alpha_k \mu}(v_k - x_k). 
        \end{align*}
    \end{observation}
        To see (Q1), observe 
        \begin{align*}
            y_k - v_k &= 
            \frac{\hat \gamma}{\alpha_k \gamma}(x_k - y_k)
            \\
            \iff 
            -(\alpha_k \gamma\hat \gamma^{-1} + 1)y_k
            &= 
            - \alpha_k \gamma \hat \gamma^{-1}v_k - x_k
            \\
            \iff 
            y_k &= 
            \frac{
                \alpha_k \gamma \hat \gamma^{-1}v_k + x_k
            }{1 + \alpha_k \gamma \hat \gamma^{-1}}
            \\
            &=  
            \frac{\alpha_k \gamma v_k + \hat \gamma x_k}{\gamma + \alpha_k \mu}.
        \end{align*}
        On the first equality (Q1), we multiplied both side of the equation by $\gamma_{k + 1}/\alpha_k \gamma_k$ and then group $y_k$ on the LHS. 
        The last equality comes by multiplying both the numerator and denominator by $\hat \gamma$, leaving the numerator $\hat \gamma + \alpha_k \gamma = \gamma + \alpha_k \mu$. 
        We used the definition of $\hat \gamma$ here. 
        To see the second equality (Q2), consider: 
        \begin{align*}
            y_k &= (\gamma + \alpha_k \mu)^{-1}(\alpha_k \gamma v_k + \hat\gamma x_k)
            \\
            \iff
            y_k - x_k &= 
            (\gamma + \alpha_k \mu)^{-1}
            (\alpha_k \gamma_k v_k - (\gamma + \alpha_k \mu)x_k + \hat \gamma x_k)
            \\
            \iff 
            (\gamma + \alpha_k \mu)(y_k - x_k)
            &= 
            \alpha_k\gamma v_k + 
            (\hat \gamma - \gamma - \alpha_k \mu) x_k
            \\
            &= \alpha_k \gamma v_k - \alpha_k \gamma x_k 
            \\
            &= \alpha_k \gamma(v_k - x_k)
            \\
            \iff 
            y_k - x_k &= 
            \frac{\alpha_k \gamma}{\gamma + \alpha_k \mu}(v_k - x_k). 
        \end{align*}
        On the second equality that follows the second $\iff$ we just substituted $\hat\gamma = (1 - \alpha_k)\gamma + \alpha_k \mu$. 
        Therefore, we discover that $x_k, v_k, y_k$ lies on the same line.

    \begin{proposition}[Stepwise lyapunov]\label{prop:stepwise-lyapunov}\;\\
        Fix any integer $k \ge0$, suppose that $v_{k + 1}, x_{k + 1}, y_k, g_k, \gamma, \hat \gamma$ satisfies Definition \ref{def:stepwise-wapg}. 

@@

            \text{use Q1}: & =

@@

            \frac{\alpha_k^2}{\hat \gamma}\frac{\hat \gamma}{\alpha_k \gamma}(x_k - y_k)

@@

            \text{use Q2}: 

@@

            \frac{\alpha_k \gamma}{\gamma_k + \alpha_k \mu}(x_k - v_k)

@@

        Initialize any $\gamma_0 \in [0, 1]$, $(x_0, v_0)$ or equivalently: $(x_0, y_0), (y_0, v_0)$. 

@@

            \text{find }&\alpha_k \in [0, 1], \rho_k \ge 0: L\alpha_k^2 = (1 - \alpha_k)\gamma_k + \mu \alpha_k  \;\wedge \; 0 \le \rho_k \alpha_k^2 \le 1, 

@@

        The sequence $\alpha_k$ satisfies the condition that $\alpha_k \in [0, 1]$.
        In addition, if $\alpha_0 \in (0, 1)$, then the entire sequence $\alpha_k \in (0, 1)$. 
        Suppose inductively that $\alpha_{k - 1}, \rho_{k - 1}$ are given such that they satisfy $\alpha_{k -1} \in [0, 1]$ and $\rho_{k - 1} \alpha_{k - 1}^2 < 1$. 
        Solving the quadratic $L\alpha_k^2=(1 - \alpha_k)\rho_{k - 1}L\alpha_{k - 1}^2 + \mu \alpha_k$ for $\alpha_k$ yields candidates 

@@

        We have $\alpha_k \ge 0$ because: 

@@

            & \ge 0. 
        If $\alpha> 0$, we would have $4\rho_{k - 1} \alpha_{k - 1}^2 > 0$ , which allows for $> 0$ instead. 
        An upper bound can be identified by using inductive hypothesis and consdering: 

@@

@@

                \frac{\mu}{L} - 1
                \sqrt{(1 - \mu/L)^2 + 4}
            &\le \frac{1}{2}\left(
                0 + \sqrt{0 + 4}
            \right) = 1. 
        Going from the second to the third inequality, we maximized $\mu/L$ by monotonicity of the linear function and the square root function. 
        Therefore, inductively, $\alpha_k \in [0, 1]$ holds. 
        Additionally, if $\alpha_{k - 1} \in (0, 1)$, then the results can be strengthened and $\alpha_k \in (0, 1)$ as well.

@@

        Recall the stepwise convergence theorem, since the choice of $\gamma, \hat \gamma$ is arbitrary choices for the theorem, we choose $\gamma = \rho_{k - 1} L \alpha_{k - 1}^2, \hat \gamma = L \alpha_k^2$ at each step, giving us: 

>>>>

\section{Preliminaries}
    Throughout Section 1, 2, 3, we assume the problem of types $F = f + g$ where $f: \RR \rightarrow \RR$ is $L$ Lipschitz smooth and $\mu \ge 0$ strongly convex and $g: Q \rightarrow \overline \RR$ is convex. 
    We consider optimization problem of the form
    \begin{align*}
        \min_x \left\lbrace
            F(x) \defeq f(x) + g(x)
        \right\rbrace. 
    \end{align*}
    To start we define the following quantities
    \begin{enumerate}
        \item The proximal gradient model function: 
        $$
            \widetilde{\mathcal M}^{L^{-1}}
            (x; y) \defeq
            g(x) + f(y) + \langle \nabla f(y), x - y\rangle 
            + \frac{L}{2}\Vert x - y\Vert^2. 
        $$
        \item The proximal point model function: 
        $$
            \mathcal M^{L^{-1}}(x; y) := F(x) + \frac{L}{2}\Vert x - y\Vert^2
        $$
        \item The proximal gradient operator $T_L = [I + L^{-1}\partial g]\circ [I - L^{-1}\nabla f]$. Since $L$ is fixed throughout, the subscript may be omitted and become $T$ instead to make the notations simpler. 
    \end{enumerate}
    To state the following lemma, we define the Bregman divergence of $f$ to be 
    \begin{align*}
        D_f(x, y): \RR^n \times \RR^n \rightarrow \RR 
        \defeq f(x) - f(y) - \langle \nabla f(y), x - y\rangle. 
    \end{align*}

@@

\section{Stepwise formulation of weak accelerated proximal gradient}
    In this section, we introduce the following new quantities and their notations. 
    It's there to make the notations simpler and easier to follow for the proofs that will come later. 
    The algorithm is parameterized by the sequence $y_k, x_k$. 
    We define for all $k \ge 0$ the following quantities: 
    \begin{assumption}
        Choose any integer $k\ge 0$. 
        Given $x_k, y_k, v_k$, we define the following quantities
        \begin{align}
            g_k &\defeq L(y_k - T_L y_k), 
            \label{eqn:grad-map}
            \\
            l_F(x; y_k) &\defeq F(T_Ly_k) + \langle g_k, x - y_k\rangle + \frac{1}{2L}\Vert g_k\Vert^2, 
            \label{eqn:lower-linearization}
            \\
            \epsilon_{k} &\defeq F(x_k) - l_F(x_k; y_k), 
            \label{eqn:regret}
        \end{align}
        Observe that by convexity of $F$, $\epsilon_k \ge 0$ always. 
        To see, consider the last inequality in the remark of Lemma \ref{thm:prox-grad-ineq} with $y = y_k, x = x_k$: 
        \begin{align*}
            F(x_k) - F(T_Ly_k)
            - \langle L(y_k - T_Ly_k),x_k - y_k \rangle
            - \frac{L}{2}\Vert y_k - T_Ly_k\Vert^2
            - \frac{\mu}{2}\Vert x_k - y_k\Vert^2
            &\ge 0
            \\
            \implies 
            F(x_k) - F(T_Ly_k)
            - \langle g_k,x_k - y_k \rangle
            - \frac{1}{2L}\Vert g_k\Vert^2
            &\ge 0. 
        \end{align*}
    \end{assumption}

    \begin{definition}[Stepwise weak accelerated proximal gradient]\label{def:stepwise-wapg}\;\\
        Let $0 \le \mu \le L$ be the strong convexity and Lipschitz smoothness parameter of $f$. 
        Given iterates $(v_k, x_k)$, or equivalently $(y_k, x_k)$, $(x_k, y_k)$, any $\alpha_k > 0, \gamma > 0$, the algorithm generates scalar $\hat \gamma$, and vectors $y_k, v_{k + 1}, x_{k + 1}$ by equalities: 
        \begin{align*}
            \hat \gamma &:= (1 - \alpha_k)\gamma + \mu \alpha_k, 
            \\
            y_k &= 
            (\gamma + \alpha_k \mu)^{-1}(\alpha_k \gamma v_k + \hat\gamma x_k), 
            \\
            g_k &= \mathcal G_L y_k, 
            \\
            v_{k + 1} &= \hat\gamma^{-1}
            (\gamma(1 - \alpha_k) v_k - \alpha_k g_k + \mu \alpha_k y_k), 
            \\
            x_{k + 1} &= T_L y_k. 
        \end{align*}
    \end{definition}
    \begin{observation}\label{obs:stepwise-wapg}
        Let's observe Definition \ref{def:stepwise-wapg} closely. 
        We make two crucial observations here. 
        We have $x_k, y_k, v_k$ lie on the same line because and they have the following equivalent equalities: 
        \begin{align*}
            (Q1): 
            y_k - v_k &= 
            \frac{\hat \gamma}{\alpha_k \gamma}(x_k - y_k),
            \\
            (Q2): 
            y_k - x_k &= 
            \frac{\alpha_k \gamma}{\gamma + \alpha_k \mu}(v_k - x_k). 
        \end{align*}
    \end{observation}
        To see (Q1), observe 
        \begin{align*}
            y_k - v_k &= 
            \frac{\hat \gamma}{\alpha_k \gamma}(x_k - y_k)
            \\
            \iff 
            -(\alpha_k \gamma\hat \gamma^{-1} + 1)y_k
            &= 
            - \alpha_k \gamma \hat \gamma^{-1}v_k - x_k
            \\
            \iff 
            y_k &= 
            \frac{
                \alpha_k \gamma \hat \gamma^{-1}v_k + x_k
            }{1 + \alpha_k \gamma \hat \gamma^{-1}}
            \\
            &=  
            \frac{\alpha_k \gamma v_k + \hat \gamma x_k}{\gamma + \alpha_k \mu}.
        \end{align*}
        On the first equality (Q1), we multiplied both side of the equation by $\hat\gamma/\alpha_k \gamma$ and then group $y_k$ on the LHS. 
        The last equality comes by multiplying both the numerator and denominator by $\hat \gamma$, leaving the numerator $\hat \gamma + \alpha_k \gamma = \gamma + \alpha_k \mu$. 
        We used the definition of $\hat \gamma$ here. 
        To see the second equality (Q2), consider: 
        \begin{align*}
            y_k &= (\gamma + \alpha_k \mu)^{-1}(\alpha_k \gamma v_k + \hat\gamma x_k)
            \\
            \iff
            y_k - x_k &= 
            (\gamma + \alpha_k \mu)^{-1}
            (\alpha_k \gamma_k v_k - (\gamma + \alpha_k \mu)x_k + \hat \gamma x_k)
            \\
            \iff 
            (\gamma + \alpha_k \mu)(y_k - x_k)
            &= 
            \alpha_k\gamma v_k + 
            (\hat \gamma - \gamma - \alpha_k \mu) x_k
            \\
            &= \alpha_k \gamma v_k - \alpha_k \gamma x_k 
            \\
            &= \alpha_k \gamma(v_k - x_k)
            \\
            \iff 
            y_k - x_k &= 
            \frac{\alpha_k \gamma}{\gamma + \alpha_k \mu}(v_k - x_k). 
        \end{align*}
        On the second equality that follows the second $\iff$ we just substituted $\hat\gamma = (1 - \alpha_k)\gamma + \alpha_k \mu$. 
        Therefore, we discover that $x_k, v_k, y_k$ lies on the same line because (Q1) indicates $y_k - v_k$ parallels to $-(y_k - x_k)$ and both vector shares the same head which anchors at $y_k$.

    \begin{proposition}[Stepwise Lyapunov]\label{prop:stepwise-lyapunov}\;\\
        Fix any integer $k \ge0$, assume $v_{k + 1}, x_{k + 1}, y_k, g_k, \hat \gamma$ are produced by Definition \ref{def:stepwise-wapg}. 

@@

            \text{use Q1 here } & =

@@

            \frac{\alpha_k^2\mu}{\hat \gamma}\frac{\hat \gamma}{\alpha_k \gamma}(x_k - y_k)

@@

            \text{use Q2 here}

@@

            \frac{\alpha_k \gamma}{\gamma + \alpha_k \mu}(x_k - v_k)

@@

        Initialize any $\gamma_0 \in (0, 1)$, $(x_0, v_0)$ or equivalently: $(x_0, y_0), (y_0, v_0)$. 

@@

            \text{find }&\alpha_k \in (0, 1), \rho_k > 0: L\alpha_k^2 = (1 - \alpha_k)\gamma_k + \mu \alpha_k  \;\wedge \; 0 < \rho_k \alpha_k^2 < 1, 

@@

        If $\alpha_0 \in (0, 1)$, then the entire sequence $\alpha_k \in (0, 1)$. 
        Suppose inductively that $\alpha_{k - 1}, \rho_{k - 1}$ are given such that they satisfy $\alpha_{k -1} \in (0, 1)$ and $0 < \rho_{k - 1} \alpha_{k - 1}^2 < 1$. 
        Solving the quadratic $L\alpha_k^2=(1 - \alpha_k)\rho_{k - 1}L\alpha_{k - 1}^2 + \mu \alpha_k$ for $\alpha_k$ yields candidates. 

@@

        We have $\alpha_k > 0$ because: 

@@

            \right) + \alpha_{k - 1}\sqrt{\rho_{k - 1}}
            & > 0. 
        On the last strict inequality we used the fact that $\rho_{k - 1}> 0, \alpha_{k - 1} > 0$. 
        An upper bound can be identified by using inductive hypothesis and considering: 

@@

            &<

@@

                \frac{\mu}{L} + 
                \sup_{x \in (0, 1)}
                \left\lbrace
                    -x + \sqrt{(x - \mu/L)^2 + x}
                \right\rbrace
            & \le \frac{1}{2}\left(
                \mu/L + \max\left(\mu/L, -1 + \sqrt{(1 - \mu/L)^2 + 1}\right)
            \right) \le 1. 
        Going to the first inequality, we used $\rho_{k - 1} \alpha_{k - 1}^2 < 1$ to get the strict inequality. 
        Going from the second to the third inequality, we maximized $\mu/L$ by monotone increasing of linear function and the square root function. 
        Therefore, inductively, $\alpha_k \in (0, 1)$ holds. 
        However, the limit of $\alpha_k$ could be 1. 

@@

        Recall the stepwise convergence theorem, since the choice of $\gamma, \hat \gamma$ is arbitrary choices for the theorem, we choose $\gamma = \rho_{k - 1} L \alpha_{k - 1}^2 > 0, \hat \gamma = L \alpha_k^2 > 0$ at each step. 
        Because $\alpha_k \in(0, 1)$ by Observation \ref{obs:r-wapg-observation-1}, we can use Proposition \ref{prop:stepwise-lyapunov}: 
